Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-13 WSSSPE1_report/WSSSPE1_summary_report.tex
Mainly Section 3.2: Keynote 2. Better flow.

Smith's talk started with an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask,'' and realized that while it was persistent, there was no
way/practice of sharing this with the data. Consequently many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

He asked, of this practice, ``why didn't we know better?'' and argued
this was because we were taught to focus on immediate research
outcomes and not on continuously improving and building on tools for
research. He then asked, later in our working lives, when we do know
better, why don't we act any different, and argued this was due to
incentives and their lack: only the immediate products of research,
not the software, are valued.  He referenced Victoria Stodden's talk
at OKCon~\cite{okcon-stodden-talk}
which he said argued these points well.

He cited the EMCEE Markov Chain Monte Carlo code~\cite{emcee-web} by
Dan Foreman-Mackey and contributors as good examples of scientific
projects using GitHub.

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant, argued that with regard to reusable
software, ``we should just start doing it.'' Smith noted that
documentation should be ``treated as a first class entity.''  He noted
that some software fields have excellent cultures of module reuse,
such as RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively ultra-low-friction
collaboration through these repositories. This has not happened in
highly numerical, compiled language scientific software.

>>>>

Smith's talk started with an example from data reduction in Astronomy, where he
needed to remove interfering effects from the device. He built a ``bad
pixel mask,'' and realized that while it was persistent, there was no
way or practice of sharing this with the data. Consequently many researchers
repeated the same calculations. He estimated that 13 person-years were
wasted in this redundant calculation.

``Why didn't we do better?''  Smith asked of this practice. He argued
this was because we were taught to focus on immediate research
outcomes and not on continuously improving and building on tools for
research. He then asked,  when we do know
better, why we do not act any different. He argued that it was due to
incentives and their lack: only the immediate products of research,
not the software, are valued.  He referenced Victoria Stodden's talk
at OKCon~\cite{okcon-stodden-talk}
which he said argued these points well.

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant, argued that with regard to reusable
software, ``we should just start doing it.'' Smith put forward that
documentation should be ``treated as a first class entity.''  He noted
that the open source community has excellent cultures of code reuse, for example,
RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively low-friction
collaboration through the use of repositories. This has not happened in
highly numerical, compiled language scientific software.
An exception he cited as a good example of scientific
projects using GitHub is the EMCEE Markov Chain Monte Carlo project~\cite{emcee-web} by
Dan Foreman-Mackey and contributors.
